Family Fortunes: A Quantitative Analysis of the Early Iron Age Cemeteries at Knossos, Crete

### 2019-04-03 analysis/paper/paper.Rmd
Writing up

colnames(aorist_all_tombs)<-c("aoristic_density")
dates<-seq(from = 610, to = 1100, by = 10)
dates<-rev(dates)
row.names(aorist_all_tombs)<-dates
aorist_all_tombs<-data.frame(aorist_all_tombs)
setDT(aorist_all_tombs, keep.rownames = "decade")
aorist_all_tombs$decade<-as.numeric(aorist_all_tombs$decade)

ggplot(aorist_all_tombs,aes(decade,aoristic_density, fill="Aoristic Sum"))+
  geom_col()+theme_bw()+theme(legend.title=element_blank(),
      text=element_text(family="Garamond", size=12))+
          labs(x="Date (Years BC)", y="Aoristic Sum")+
              scale_fill_brewer(palette = "Set3")+ 
                theme(legend.position="none")+
                  scale_y_continuous(expand=c(0,0),limits=c(0,225))+
                    scale_x_reverse(expand = c(0, 0),limits=c(1100,600))

@@

colnames(aorist_KNC_pithoi)<-c("aoristic_density")
row.names(aorist_KNC_pithoi)<-dates
aorist_KNC_pithoi<-data.frame(aorist_KNC_pithoi)
setDT(aorist_KNC_pithoi, keep.rownames = "decade")
aorist_KNC_pithoi$decade<-as.numeric(aorist_KNC_pithoi$decade)

@@

colnames(aorist_FOR_pithoi)<-c("aoristic_density")
row.names(aorist_FOR_pithoi)<-dates
aorist_FOR_pithoi<-data.frame(aorist_FOR_pithoi)
setDT(aorist_FOR_pithoi, keep.rownames = "decade")
aorist_FOR_pithoi$decade<-as.numeric(aorist_FOR_pithoi$decade)

@@

colnames(aorist_all_pithoi)<-c("aoristic_density")

@@

aorist_pithoi_plot<-ggplot(NULL)+
  geom_col(data=aorist_all_pithoi,aes(decade,aoristic_density,fill=" All tombs  "))+
    geom_col(data=aorist_KNC_pithoi,aes(decade,aoristic_density,fill=" KNC  "))+
      geom_col(data=aorist_FOR_pithoi,aes(decade,aoristic_density, fill=" Fortetsa  "))+
        theme_bw()+theme(legend.title=element_blank(),
          text=element_text(family="Garamond", size=12))+
            labs(x="Date (Years BC)", y="Aoristic Sum (Pithoi)")+
              scale_fill_brewer(palette = "Set2")+ 
                  scale_y_continuous(expand=c(0,0),limits=c(0,45))+
                    scale_x_reverse(expand = c(0, 0),limits=c(1100,600))+
                      theme(plot.margin=unit(c(0.1,1,0.1,0.1),"cm"))

@@

colnames(aorist_KNC)<-c("aoristic_density")
row.names(aorist_KNC)<-dates
aorist_KNC<-data.frame(aorist_KNC)
setDT(aorist_KNC, keep.rownames = "decade")
aorist_KNC$decade<-as.numeric(aorist_KNC$decade)

@@

colnames(aorist_FOR)<-c("aoristic_density")
row.names(aorist_FOR)<-dates
aorist_FOR<-data.frame(aorist_FOR)
setDT(aorist_FOR, keep.rownames = "decade")
aorist_FOR$decade<-as.numeric(aorist_FOR$decade)

aorist_pots_plot<-ggplot(NULL)+
  geom_col(data=aorist_all_tombs,aes(decade,aoristic_density,fill=" All tombs  "))+
    geom_col(data=aorist_KNC,aes(decade,aoristic_density,fill=" KNC  "))+
      geom_col(data=aorist_FOR,aes(decade,aoristic_density, fill=" Fortetsa  "))+
        theme_bw()+theme(legend.title=element_blank(),
          text=element_text(family="Garamond", size=12))+
            labs(x="Date (Years BC)", y="Aoristic Sum (All Vessels)")+
              scale_fill_brewer(palette = "Set2")+ 
                  scale_y_continuous(expand=c(0,0),limits=c(0,225))+
                    scale_x_reverse(expand = c(0, 0),limits=c(1100,600))+
                      theme(plot.margin=unit(c(0.1,1,0.1,0.1),"cm"))

@@

The near total absence of pithoi from the 11th and early 10th centuries reflects the delayed onset of cremation burial, with inhumations representing a significant proportion of earlier interments. Nonetheless, the clear inference would appear to be that, at both cemeteries, the late 8th and early 7th centuries witnessed an unprecedented level of burial activity, something which makes the sudden and total drop of chamber-tomb burials around 630 BC even more remarkable.

```{r aoristknc}
KNC<-subset(tombs,tombs$cem_condensed=="NC")

colnames(aorist_KNC)<-c("aoristic_density")
row.names(aorist_KNC)<-dates
aorist_KNC<-data.frame(aorist_KNC)
setDT(aorist_KNC, keep.rownames = "decade")
aorist_KNC$decade<-as.numeric(aorist_KNC$decade)

colnames(aorist_FOR)<-c("aoristic_density")
row.names(aorist_FOR)<-dates
aorist_FOR<-data.frame(aorist_FOR)
setDT(aorist_FOR, keep.rownames = "decade")
aorist_FOR$decade<-as.numeric(aorist_FOR$decade)

ggplot(NULL)+geom_col(data=aorist_all_tombs,aes(decade,aoristic_density,fill="All tombs"))+
  geom_col(data=aorist_KNC,aes(decade,aoristic_density,fill="KNC"))+
  geom_col(data=aorist_FOR,aes(decade,aoristic_density, fill="Fortetsa"))+
      xlim(1100,630)+theme_minimal()+
        theme(legend.title=element_blank(),
          text=element_text(family="Garamond", size=12))+
              labs(x="Date (Years BC)", y="Aoristic Density")+
                 scale_fill_brewer(palette = "Set2")

@@

aorist_pot_medians$aorist_sums<-aorist_all_tombs$aoristic_density
aorist_pot_medians$aorist_median_adjust<-aorist_pot_medians$aorist_sums/aorist_pot_medians$aoristic_median
View(aorist_pot_medians)

@@

colnames(aorist_pithoi)<-c("aoristic_density")
row.names(aorist_pithoi)<-dates
aorist_pithoi<-data.frame(aorist_pithoi)
setDT(aorist_pithoi, keep.rownames = "decade")
aorist_pithoi$decade<-as.numeric(aorist_pithoi$decade)

@@

colnames(aorist_all_urns)<-c("aoristic_density")
row.names(aorist_all_urns)<-dates
aorist_all_urns<-data.frame(aorist_all_urns)
setDT(aorist_all_urns, keep.rownames = "decade")
aorist_all_urns$decade<-as.numeric(aorist_all_urns$decade)

ggplot(NULL)+geom_col(data=aorist_all_urns,aes(decade,aoristic_density,fill="All urns"))+
  geom_col(data=aorist_pithoi,aes(decade,aoristic_density,fill="Pithoi"))+
      xlim(1100,630)+theme_minimal()+
        theme(legend.title=element_blank(),
          text=element_text(family="Garamond", size=12))+
              labs(x="Date (Years BC)", y="Aoristic Density")+
                 scale_fill_brewer(palette = "Set2")

@@

aorist_by_tomb$aorist_by_tomb<-aorist_by_tomb$aoristic_density/aorist_by_tomb$no_tombs

@@

aorist_tombs_9th<-colSums(tombs_9th[157:206])
aorist_tombs_9th<-data.frame(aorist_tombs_9th)
colnames(aorist_tombs_9th)<-c("aoristic_density")
row.names(aorist_tombs_9th)<-dates
aorist_tombs_9th<-data.frame(aorist_tombs_9th)
setDT(aorist_tombs_9th, keep.rownames = "decade")
aorist_tombs_9th$decade<-as.numeric(aorist_tombs_9th$decade)

aorist_by_tomb_9th<-merge(aorist_tombs_9th,no_tombs_9th)
aorist_by_tomb_9th$aorist_by_tomb<-aorist_by_tomb_9th$aoristic_density/aorist_by_tomb_9th$no_tombs

@@

tombs$start_period<-factor(tombs$start_period,levels=c("SM", "EPG", "MPG","PG", "LPG", "PGB", "EG", "MG","G", "LG", "EO","O", "LO"))

@@

aorist_tombs_top<-colSums(tombs_top_12[157:206])
aorist_tombs_top<-data.frame(aorist_tombs_top)
colnames(aorist_tombs_top)<-c("aoristic_density")
row.names(aorist_tombs_top)<-dates
aorist_tombs_top<-data.frame(aorist_tombs_top)
setDT(aorist_tombs_top, keep.rownames = "decade")
aorist_tombs_top$decade<-as.numeric(aorist_tombs_top$decade)

aorist_by_tomb_top<-merge(aorist_tombs_top,no_tombs_top)
aorist_by_tomb_top$aorist_by_tomb<-aorist_by_tomb_top$aoristic_density/aorist_by_tomb_top$no_tombs

>>>>

colnames(aorist_all_tombs)<-c("aoristic_sum")
dates<-seq(from = 610, to = 1100, by = 10)
dates<-rev(dates)
row.names(aorist_all_tombs)<-dates
aorist_all_tombs<-data.frame(aorist_all_tombs)
setDT(aorist_all_tombs, keep.rownames = "decade")
aorist_all_tombs$decade<-as.numeric(aorist_all_tombs$decade)

ggplot(aorist_all_tombs,aes(decade,aoristic_sum, fill="Aoristic Sum"))+
  geom_col()+theme_bw()+theme(legend.title=element_blank(),
      text=element_text(family="Garamond", size=12))+
          labs(x="Date (Years BC)", y="Aoristic Sum")+
              scale_fill_brewer(palette = "Set3")+ 
                theme(legend.position="none")+
                  scale_y_continuous(expand=c(0,0),limits=c(0,225))+
                    scale_x_reverse(expand = c(0, 0),limits=c(1100,600))

@@

colnames(aorist_KNC_pithoi)<-c("aoristic_sum")
row.names(aorist_KNC_pithoi)<-dates
aorist_KNC_pithoi<-data.frame(aorist_KNC_pithoi)
setDT(aorist_KNC_pithoi, keep.rownames = "decade")
aorist_KNC_pithoi$decade<-as.numeric(aorist_KNC_pithoi$decade)

@@

colnames(aorist_FOR_pithoi)<-c("aoristic_sum")
row.names(aorist_FOR_pithoi)<-dates
aorist_FOR_pithoi<-data.frame(aorist_FOR_pithoi)
setDT(aorist_FOR_pithoi, keep.rownames = "decade")
aorist_FOR_pithoi$decade<-as.numeric(aorist_FOR_pithoi$decade)

@@

colnames(aorist_all_pithoi)<-c("aoristic_sum")

@@

aorist_pithoi_plot<-ggplot(NULL)+
  geom_col(data=aorist_all_pithoi,aes(decade,aoristic_sum,fill=" All tombs  "))+
    geom_col(data=aorist_KNC_pithoi,aes(decade,aoristic_sum,fill=" KNC  "))+
      geom_col(data=aorist_FOR_pithoi,aes(decade,aoristic_sum, fill=" Fortetsa  "))+
        theme_bw()+theme(legend.title=element_blank(),
          text=element_text(family="Garamond", size=12))+
            labs(x="Date (Years BC)", y="Aoristic Sum (Pithoi)")+
              scale_fill_brewer(palette = "Set2")+ 
                  scale_y_continuous(expand=c(0,0),limits=c(0,45))+
                    scale_x_reverse(expand = c(0, 0),limits=c(1100,600))+
                      theme(plot.margin=unit(c(0.1,1,0.1,0.1),"cm"))

@@

colnames(aorist_KNC)<-c("aoristic_sum")
row.names(aorist_KNC)<-dates
aorist_KNC<-data.frame(aorist_KNC)
setDT(aorist_KNC, keep.rownames = "decade")
aorist_KNC$decade<-as.numeric(aorist_KNC$decade)

@@

colnames(aorist_FOR)<-c("aoristic_sum")
row.names(aorist_FOR)<-dates
aorist_FOR<-data.frame(aorist_FOR)
setDT(aorist_FOR, keep.rownames = "decade")
aorist_FOR$decade<-as.numeric(aorist_FOR$decade)

aorist_pots_plot<-ggplot(NULL)+
  geom_col(data=aorist_all_tombs,aes(decade,aoristic_sum,fill=" All tombs  "))+
    geom_col(data=aorist_KNC,aes(decade,aoristic_sum,fill=" KNC  "))+
      geom_col(data=aorist_FOR,aes(decade,aoristic_sum, fill=" Fortetsa  "))+
        theme_bw()+theme(legend.title=element_blank(),
          text=element_text(family="Garamond", size=12))+
            labs(x="Date (Years BC)", y="Aoristic Sum (All Vessels)")+
              scale_fill_brewer(palette = "Set2")+ 
                  scale_y_continuous(expand=c(0,0),limits=c(0,225))+
                    scale_x_reverse(expand = c(0, 0),limits=c(1100,600))+
                      theme(plot.margin=unit(c(0.1,1,0.1,0.1),"cm"))

@@

The near total absence of pithoi from the 11th and early 10th centuries reflects the delayed onset of cremation burial, with inhumations representing a significant proportion of earlier interments. Nonetheless, the clear inference would appear to be that, at both cemeteries, the late 8th and early 7th centuries witnessed an unprecedented level of burial activity, something which makes the sudden and total decline of chamber-tomb burials around 630 BC even more notable.

```{r tombconstruction}

tomb_dates_aorist<-read.csv("../data/derived_data/tomb_dates_aorist.csv",header=T)
tomb_dates_aorist<-subset(tomb_dates_aorist,tomb_dates_aorist$cem_condensed=="NC"|
                            tomb_dates_aorist$cem_condensed=="FOR")

aorist_tomb_dates<-colSums(tomb_dates_aorist[16:65])
aorist_tomb_dates<-data.frame(aorist_tomb_dates)

colnames(aorist_tomb_dates)<-c("aoristic_sum")
row.names(aorist_tomb_dates)<-dates
aorist_tomb_dates<-data.frame(aorist_tomb_dates)
setDT(aorist_tomb_dates, keep.rownames = "decade")
aorist_tomb_dates$decade<-as.numeric(aorist_tomb_dates$decade)

aorist_tomb_values<-tomb_dates_aorist[16:65]
aorist_tomb_matrix<-data.matrix(aorist_tomb_values)
aorist_tomb_matrix[aorist_tomb_matrix == 0] <- NA
aorist_tomb_dates$medians<-apply(aorist_pot_matrix,2,median,na.rm=TRUE)

aorist_tomb_dates$median_adjust<-aorist_tomb_dates$aoristic_sum/aorist_tomb_dates$medians

ggplot(aorist_tomb_dates,aes(decade,median_adjust))+geom_col()+
  theme_bw()+theme(legend.title=element_blank(),
          text=element_text(family="Garamond", size=12))+
            labs(x="Date (Years BC)", y="Aoristic Sum (All Vessels)")+
              scale_fill_brewer(palette = "Set2")+ 
                  scale_y_continuous(expand=c(0,0),limits=c(0,65))+
                    scale_x_reverse(expand = c(0, 0),limits=c(1100,600))+
                      theme(plot.margin=unit(c(0.1,1,0.1,0.1),"cm"))

aorist_tomb_periods <-colSums(tomb_dates_aorist[5:14])
aorist_tomb_periods<-data.frame(aorist_tomb_periods)

colnames(aorist_tomb_periods)<-c("aoristic_sum")
row.names(aorist_tomb_periods)<-c("SM", "EPG", "MPG", "LPG", "PGB", "EG", "MG", "LG", "EO", "LO")
aorist_tomb_periods<-data.frame(aorist_tomb_periods)
setDT(aorist_tomb_periods, keep.rownames = "period")
aorist_tomb_periods$period<-factor(aorist_tomb_periods$period,levels=c("SM", "EPG", "MPG", "LPG", "PGB", "EG", "MG", "LG", "EO", "LO"))

ggplot(aorist_tomb_periods,aes(period,aoristic_sum, fill="Period"))+geom_col()+
  theme_bw()+theme(legend.title=element_blank(),
          text=element_text(family="Garamond", size=12))+
            labs(x="Period", y="Aoristic Sum")+
              scale_fill_brewer(palette = "Set2")+ 
                 scale_y_continuous(expand=c(0,0),limits=c(0,30))+
                    theme(legend.position="none")

@@

aorist_pot_medians$aoristic_sums<-aorist_all_tombs$aoristic_sum
aorist_pot_medians$aorist_median_adjust<-aorist_pot_medians$aoristic_sums/aorist_pot_medians$aoristic_median

@@

colnames(aorist_pithoi)<-c("aoristic_sum")
row.names(aorist_pithoi)<-dates
aorist_pithoi<-data.frame(aorist_pithoi)
setDT(aorist_pithoi, keep.rownames = "decade")
aorist_pithoi$decade<-as.numeric(aorist_pithoi$decade)

@@

colnames(aorist_all_urns)<-c("aoristic_sum")
row.names(aorist_all_urns)<-dates
aorist_all_urns<-data.frame(aorist_all_urns)
setDT(aorist_all_urns, keep.rownames = "decade")
aorist_all_urns$decade<-as.numeric(aorist_all_urns$decade)

ggplot(NULL)+geom_col(data=aorist_all_urns,aes(decade,aoristic_sum,fill="All urns"))+
  geom_col(data=aorist_pithoi,aes(decade,aoristic_sum,fill="Pithoi"))+
      xlim(1100,630)+theme_minimal()+
        theme(legend.title=element_blank(),
          text=element_text(family="Garamond", size=12))+
              labs(x="Date (Years BC)", y="Aoristic Density")+
                 scale_fill_brewer(palette = "Set2")

@@

aorist_by_tomb$aorist_by_tomb<-aorist_by_tomb$aoristic_sum/aorist_by_tomb$no_tombs

@@

aorist_tombs_9th<-colSums(tombs_9th[157:206])
aorist_tombs_9th<-data.frame(aorist_tombs_9th)
colnames(aorist_tombs_9th)<-c("aoristic_sum")
row.names(aorist_tombs_9th)<-dates
aorist_tombs_9th<-data.frame(aorist_tombs_9th)
setDT(aorist_tombs_9th, keep.rownames = "decade")
aorist_tombs_9th$decade<-as.numeric(aorist_tombs_9th$decade)

aorist_by_tomb_9th<-merge(aorist_tombs_9th,no_tombs_9th)
aorist_by_tomb_9th$aorist_by_tomb<-aorist_by_tomb_9th$aoristic_sum/aorist_by_tomb_9th$no_tombs

@@

@@

aorist_tombs_top<-colSums(tombs_top_12[157:206])
aorist_tombs_top<-data.frame(aorist_tombs_top)
colnames(aorist_tombs_top)<-c("aoristic_sum")
row.names(aorist_tombs_top)<-dates
aorist_tombs_top<-data.frame(aorist_tombs_top)
setDT(aorist_tombs_top, keep.rownames = "decade")
aorist_tombs_top$decade<-as.numeric(aorist_tombs_top$decade)

aorist_by_tomb_top<-merge(aorist_tombs_top,no_tombs_top)
aorist_by_tomb_top$aorist_by_tomb<-aorist_by_tomb_top$aoristic_sum/aorist_by_tomb_top$no_tombs
